---
title: "Week 1 Answers"
author: "Brett Lambert"
date: "8/22/2024"
output: html_document
---

# Basic R Practice Answers

## Exercises

For the following exercises, please create code chunks and type out all your code after each question or task.

1. Describe your previous experience with R. 
```{r}
# I have experience using R for data manipulation, cleaning, multivariate analyses like PCA, population genetic analyses, and phylogenetic comparative methods. I have read most of R4DS and also written a script to pull genbank records given a CSV of accession numbers.
```


1. What skills do you hope to gain from this course? What reservations do you have regarding this course? 
```{r}
#I hope to gain some new insights on application of methods to real data and also hoping to understand more clearly which aspects of R I should focus on to become more marketable to industry careers.
```

1. Write a short paragraph on your research using 4 R Markdown text formattings. 

#Species delimitation within _Asarum canadense_

My research focuses on a species of plant that exhibits a wide range of morphology; previously leading to the recognition of **3 species**. This system is currently recognized as 1 species, however and we want to investigate the different morphotypes with morhpological and genetic data to determine if these are distinct species or one phenotypically plastic species. This research is currently ~~making me question my life choices~~ going well and hopefully will be finished within a year.


## Search for the filter function. 

```{r}
#?filter()
```



## Search for the dplyr package and look up summarize within this package. 

```{r}
#?dplyr::summarise()
```



## Set your working directory to your pictures folder. 

```{r}
#Don't really want to change directory. I use cloud-based folders so it is not straightforawrd.
library(tidyverse)
getwd()
#?setwd()
#setwd("C:/Users/brettl71/Pictures")

#setwd("Users/brettl71/Library/CloudStorage/OneDrive-TheOhioStateUniversity/OSU_Courses_3_22_24/AU_24/Data_Tidying_Seminar/Week_1
```



## Import a dataset or table.

```{r}
ex <- read.csv("megaGymDataset.csv")
```



## What is the structure of your data? What are your column headers? Change the class of your data for this R session.

```{r}
View(ex)
str(ex)
as.numeric(ex$Rating)
```



## Create an object that has a minimum of 10 pieces of data. What is the structure of this data? 

```{r}
obj <- c(1,33,4,12,9,0,35,224,198,86)
str(obj)
```



## Use the R dataset "chickwts". What are the dimensions of this dataset? Use attach to create a simple boxplot of weight by feed type. 

```{r}
chc <- chickwts
dim(chc)
attach(chc)
boxplot(weight ~ feed)
detach(chc)
```



## Add a column of data called "DPH" (days post hatch) to the chickwts dataset. There are multiple ways to add a column to a dataset. While we did not go over adding a column explicitly, use this question as a way to learn the help functions and increase your savvy with googling R questions. Save this dataset to your desktop. 

```{r}
chc$DPH <- 0
chc

chc2 <- chc |>
  mutate("DPH" = 0)
chc2

?mutate
```

## Attempt to knit the document to an .html or other file format once completed. Turn in both the R Markdown (.rmd) and knitted file if you were successful. 