---
title: "Week 2 In Class Assignment Intro to Data Wrangling"
author: "S.A. Valentine"
date: "2024-08-20"
output: html_document
---

You will use the Stream solutes data to answer the following questions. You will use the tools we have covered so far within `dplyr`. Please annotate your code, provide brief reasoning as to why you used specific functions to complete the tasks, and provide a brief interpretation of the result of the code chunk.

Task 1: Read in data. How many columns and rows are there? 

What is the mode of each variable in the dataset? 

```{r Task 1}
library(tidyverse)
streams <- read_csv("streamsolutes.csv")
dim(streams)
summary(streams)
```


Task 2: How many sites are in the dataset? How many solutes and what are the solutes?

```{r Task 2}
unique(streams$Site)

length(unique(streams$Solute))
print(unique(streams$Solute))
```

Task 3 Rename the Q_Ls column to "Discharge (Ls)", then keep only the new discharge column and Site, Solute, and mgL. Save these updated data as a new object.

```{r Task 3}
new_streams <- rename(streams, Discharge_Ls = Q_Ls) 
new_streams_1 <- select(new_streams, c("Discharge_Ls", "Site", "Solute", "mgL"))

#streams_BBWM$Date <- as.numeric(streams_BBWM$Date)
```


Task 4: Filter the new dataset for site "BBWM". Make a quick plot of date and discharge.


```{r Task 4}
streams_BBWM <- filter(new_streams, Site == "BBWM")
ggplot(data = streams_BBWM, mapping = aes(x = Date, y = Discharge_Ls)) +
                 geom_point()
?ggplot()
```

Task 5: Filter the original dataset for site "TLW" and solute "Ca" and save it as a new object. Calculate the minimum, maximum, mode, and mean of the Ca values.

```{r}
TLW_Ca <- filter(streams, Site == "TLW" & Solute == "Ca")
```



Task 6: Create a new column called "Discharge cfs" (cubic feet per second). The values should be converted from Q_Ls (liters per second) from the original dataset. What is the mean, maximum, and minimum values of Discharge (cfs) for the year 1998 in site ELA?  

```{r Task 6}
new_col <- mutate(streams, Discharge_cfs = Q_Ls/28.32)
str(new_col)
test <- new_col |>
 filter(between(Date, as.Date('1998-01-20'), as.Date('1998-12-31')))
```


Task 7: Where and when did the highest NH4 value occur? 

```{r Task 7}
View(test)
NH4 <- filter(test, Solute == "NH4")
str(NH4)
names(NH4)
Ordered_NH4<- NH4 |>
  arrange(desc(Discharge_cfs))

```