---
title: "Week 3 Assignment Data Wrangling with the Pipe"
author: "Brett Lambert"
date: "2024-09-04"
output: html_document
---

You will use the tools we have covered so far within `dplyr` and synthesis the information by using the pipe. You will use the Gillnet catches from 1930-1932 in Lake Michigan data to answer the following questions. Please annotate your code, provide brief reasoning as to why you used specific functions to complete the tasks, and provide a brief interpretation of the result of the code chunk.



Task 1: What is the total number of fish by species in each GN_EFFORT_ID? 

What is the total number of individuals of species 307 for each gillnet ID? 

Which gillnet effort has the greatest catch of species 307 (use arrange function)?

```{r Task 1}
#total number of fish by species in each GN_EFFORT_ID


#Group by gillnet and species
#summarize by total number using summarize with n() or count
library(tidyverse)
gn_data <- read_csv("lakeM_Fulmar_Catch.csv")

grouped_df <- group_by(gn_data, GN_EFFORT_ID, SPECIES) |> summarise(total_count=sum(N))

print(unique(grouped_df$total_count))



#total number of inds for each gillnet ID for species 307
#use code from above and filter to include only records for species 307
filtered_df <- group_by(gn_data, GN_EFFORT_ID, SPECIES) |> filter(SPECIES == 307) |>
  summarise(total_count=sum(N)) 

print(unique(filtered_df$SPECIES))
#It looks like we've successfully filtered to only have 307


#Greatest catch of species 307
#Same code as above and arrange by column representing number of fish.
max_df <- group_by(gn_data, GN_EFFORT_ID, SPECIES) |> filter(SPECIES == 307) |>
  summarise(total_count=sum(N)) |>
  arrange(desc(total_count))

print(max(max_df$total_count))

#Highest count of fish for species 307 is 186
```


Task 2: Which gillnet ID(s) has/have the greatest species richness (highest diversity/number of species) and what is that richness? How many gillnet IDs have a species richness of 3-7? 

```{r Task 2}
#Group by GN_EFFORT_ID and SPECIES
grouped_df_2 <- group_by(gn_data, GN_EFFORT_ID, SPECIES)
#summarize data using any function I usually use n() to get total observations for each GN, Species combination, although the summarized column is not used.
species_df <- group_by(gn_data, GN_EFFORT_ID) |>
  summarise(total_count = n_distinct(SPECIES)) |>
  arrange(desc(total_count))

#GN_Effort_ID 44260 has the highest richness at 13

print(max(species_df$total_count))

#Group by GN_EFFORT_ID
species_df_2 <- group_by(gn_data, GN_EFFORT_ID) |>
  summarise(total_count = n_distinct(SPECIES)) |>
  filter(total_count < 8 & total_count > 2) |>
  arrange(desc(total_count))

length(unique(species_df_2$GN_EFFORT_ID))
#868

```

Task 3 What is the mean, max, and min Weight (in kg) as well as sample size for each species within gillnet ID 44365? 

Hint: First, remove all rows without weight data. 

What is the mean, max, and min Weight (in kg) as well as sample size for each species across all gillnets? 

Which gillnet had the largest fish? Which species and how big was the largest fish in weight and length in cm? 

```{r Task 3}
#Task 3 What is the mean, max, and min Weight (in kg) as well as sample size for each species within gillnet ID 44365? 

#Get cleaned data for further processing. Remove NAs in weight column
weight_data <- gn_data |>
  filter(!is.na(WEIGHT)) |>
  mutate(WEIGHT = WEIGHT/1000)

#Filter for desired GN_EFFORT_ID
weight_44365 <- weight_data |>
  group_by(GN_EFFORT_ID) |>
  filter(GN_EFFORT_ID == 44365) 

#Get summary stats of weight for observations having the specific ID
summary(weight_44365$WEIGHT)

#print unique species obs and corresponding count
print(select(weight_44365, c(SPECIES, N)))




#What is the mean, max, and min Weight (in kg) as well as sample size for each species across all gillnets?

#For each species across all gillnets, group by species
species_group <- weight_data |>
  group_by(SPECIES) |>
  arrange(SPECIES, .locale = "en")
summary(species_group$WEIGHT)


species_group_1 <- weight_data |>
  group_by(SPECIES) |>
  summarise(N = n())




#Which gillnet had the largest fish? Which species and how big was the largest fish in weight and length in cm?

#Arrange by weight
by_weight <- weight_data |>
  arrange(desc(WEIGHT))

#Take out row that is the max of weight and compare to arranged df
maximum <- weight_data |>
  filter(WEIGHT == max(WEIGHT))
```


